Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-14 chi2012.tex
incremental updates
consolidation action

\begin{abstract}

\end{abstract}

@@

If PIM tools were extended to enable end-users to make effective use of the emerging ecosystem of personal data, how would this impact personal information practice? This paper presents our initial results towards answering this question through a multi-method investigation of the personal data integration problem.  We start with a small interview study examining information tasks people perform that rely on multiple sources of information, and how such needs are currently satisfied.  Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.  Second, we conduct a analysis of various popular personal data sources available today to identify barriers to effective unification of data from each.  We draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

@@

\section{Pre-studies: The Need and Challenge of Data Integration in PIM}

@@

@@

In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.  For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).  We analysed the data collected from this study using a thematic approach.  From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).  We performed a statistical analysis of the features of the available data.  In the next two sections, we describe our two pre-studies.
%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

@@

\subsubsection{Question 2: Large Social Event}
All of the participants would organise a large social event by talking to their friends, about their ideas, preferences and recommendations.  The tools they would use to acquire this information included face-to-face, phone calls, email, Skype, Facebook events.  The method chosen depended on the time scale required to organise an event, in general if there was a short time scale people would speak in person or on the phone.  Otherwise, people would set up a FaceBook event page to discuss ideas with their friends.  If their friends were not on Facebook they would email them.  All but one participant said they would not use Doddle to organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work event not social.

Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.  They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.   They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.  People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.  They considered how to get there and looked up train or bus times on the Web.  One participant organised carpools informally and at last minute via Facebook or text messages.

@@

\subsection{PS.2 - Technical Challenges of Integrating the Data}
Our investigation of integration challenges started with identifying a set of data sources to analyze.  For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.  In order to get a variety of different kinds of sources represented, we feeds pertaining to 5 topics: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.   Table The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.

We first sought out the APIs and data feeds of popular Web services and sites that were aimed at developers creating apps for end-users.  Based on statistics collected We first selected 4 different categories

@@

>>>>

\begin{abstract}
Exploratory sequential mixed method study
\end{abstract}

@@

If PIM tools were extended to enable end-users to make effective use of the emerging ecosystem of personal data, how would this impact personal information practice? This paper presents our initial results towards answering this question through a sequential exploratory mixed-method investigation of the personal data integration problem.  We start with a small interview study examining information tasks people perform that rely on multiple sources of information, and how such needs are currently satisfied.  Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.  Second, we conduct a analysis of various popular personal data sources available today to identify barriers to effective unification of data from each.  We draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

@@

\section{Pre-studies: The Need for and Challenges in realizing Ad-Hoc Data Integration in PIM Tools}

@@

% blargh. this needs to be integrated above! 
In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed-method investigation to help us design our contribution.  For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).  We analysed the data collected from this study using a thematic approach.  From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).  We performed a statistical analysis of the features of the available data.  In the next two sections, we describe our two pre-studies.

@@

%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

@@

\subsubsection{Question 2: Large Social Event}
All of the participants would organise a large social event by talking to their friends, about their ideas, preferences and recommendations.  The tools they would use to acquire this information included face-to-face, phone calls, email, Skype, Facebook events.  The method chosen depended on the time scale required to organise an event, in general if there was a short time scale people would speak in person or on the phone.  Otherwise, people would set up a Facebook event page to discuss ideas with their friends.  If their friends were not on Facebook they would email them.  All but one participant said they would not use Doodle\footnote{Doodle - \emph{www.doodle.com}} to organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work event not social.

Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.  They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.   They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews obtained through web search and sites such as tripadvisor.  People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.  They considered how to get there and looked up train or bus times on the Web.  One participant organised carpools informally and at last minute via Facebook or text messages.

@@

@@

\subsection{PS.2 - Technical Challenges of Integrating the Data}
Our investigation of integration challenges started with identifying a set of data sources to analyze.  For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.  In order to get a variety of different kinds of sources represented, based upon available choices, we first chose 5 feed \emph{categories}: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.   Then, we populated each category with the most popular feeds from the list, and took samples of data records from each.
